---
title: "INFO523_Project_SVM"
author: "Rice"
date: "2023-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(e1071)
library(ggplot2)

# Load the MASS package to access the cats dataset
library(MASS)

# Access the cats dataset
my_data <- cats

plot(my_data$Hwt, my_data$Bwt, col=as.factor(my_data$Sex), 
     main="Scatterplot of Bwt vs Hwt", 
     xlab="Heart Weight (Hwt)", ylab="Body Weight (Bwt)")
legend("topright", legend=levels(my_data$Sex), fill=1:length(levels(my_data$Sex)))



# Scale only the numeric columns. For this dataset, we'll scale Bwt and Hwt
my_data$Bwt <- scale(my_data$Bwt) # scaling the column Bwt
my_data$Hwt <- scale(my_data$Hwt) # scaling the colun Hwt

my_data$rand <- runif(nrow(my_data)) # adding a column of random values 
my_data <- my_data[order(my_data$rand),] # sort by column of random values
my_data <- my_data[,-ncol(my_data)] # deleting column of random values 

train <- my_data[1:100,]
test <- my_data[101:144,]



model_linear <- svm(Sex~., data = train, kernel="linear", C=10)
plot <- plot(model_linear,train)

m_predict_linear <- predict(model_linear,test)
table(m_predict_linear, test$Sex)

agreement <- m_predict_linear == test$Sex
table(agreement)
round(prop.table(table(agreement)),2)


model_radial <- svm(Sex~., data = train, kernel="radial", C=10, gamma=0.8)
plot <- plot(model_radial, train)
model_predict_radial <- predict(model_radial,test)
table(model_predict_radial, test$Sex)

agreement <- model_predict_radial == test$Sex
table(agreement)
round(prop.table(table(agreement)),2)



model_poly <- svm(Sex~., data = train, kernel="polynomial", degree=5, gamma=0.8)
plot <- plot(model_poly, train)
model_predict_poly <- predict(model_poly,test)
table(model_predict_poly, test$Sex)

agreement <- model_predict_poly == test$Sex
table(agreement)
round(prop.table(table(agreement)),2)




# X rep. the support vectors - data points that line on or within the margin
# O rep. non-support vectors - lie outside the margin and do not directly influence the position or orientation of the hyperplane

```

```{r}
library(kernlab)

```

```{r}
## download the letter recognition data set 

library(here)
if (!"letter-recognition.data" %in% list.files(here("data"))) {
  url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/letter-recognition/letter-recognition.data"
  download.file(url, destfile=here("data","letter-recognition.data"), mode="wb") 
}


```


```{r}

## reading in the letter recognition data set 

letters <- read.csv(here("data", "letter-recognition.data"), header = FALSE,stringsAsFactors = TRUE)

letters <- data.frame(letters)

colnames(letters) <- c("Letter","x-box","y-box","width","high","onpix","x-bar","y-bar","x2bar","y2bar","xybar","x2ybr","xy2br","x-ege","xegvy","y-ege","yegvx")

letters_train <- letters[1:16000,]
letters_test <- letters[16001:20000,]

head(letters)

```

Attribute Information:
	 1.	lettr	capital letter	(26 values from A to Z)
	 2.	x-box	horizontal position of box	(integer)
	 3.	y-box	vertical position of box	(integer)
	 4.	width	width of box			(integer)
	 5.	high 	height of box			(integer)
	 6.	onpix	total # on pixels		(integer)
	 7.	x-bar	mean x of on pixels in box	(integer)
	 8.	y-bar	mean y of on pixels in box	(integer)
	 9.	x2bar	mean x variance			(integer)
	10.	y2bar	mean y variance			(integer)
	11.	xybar	mean x y correlation		(integer)
	12.	x2ybr	mean of x * x * y		(integer)
	13.	xy2br	mean of x * y * y		(integer)
	14.	x-ege	mean edge count left to right	(integer)
	15.	xegvy	correlation of x-ege with y	(integer)
	16.	y-ege	mean edge count bottom to top	(integer)
	17.	yegvx	correlation of y-ege with x	(integer)
	
```{r}

svm_linear <- ksvm(Letter ~ ., data = letters_train, kernel = "vanilladot")

letter_predictions <- predict(svm_linear, letters_test)

table(letter_predictions, letters_test$Letter)

agreement <- letter_predictions == letters_test$Letter
table(agreement)
prop.table(table(agreement))
```

```{r}
set.seed(12345)
letter_classifier_rbf <- ksvm(Letter ~ ., data = letters_train,kernel = "rbfdot")
letter_predictions_rbf <- predict(letter_classifier_rbf, letters_test)
agreement_rbf <- letter_predictions_rbf == letters_test$Letter

table(letter_predictions_rbf, letters_test$Letter)

agreement <- letter_predictions == letters_test$Letter
table(agreement)
prop.table(table(agreement))

```

```{r}
cost_values <- c(1, seq(from = 4, to = 20, by = 2))
accuracy_values <- sapply(cost_values, function(x) {
 set.seed(12345)
 m <- ksvm(Letter ~ ., data = letters_train, kernel = "rbfdot", C = x)
 pred <- predict(m, letters_test)
 agree <- ifelse(pred == letters_test$Letter, 1, 0)
 accuracy <- sum(agree) / nrow(letters_test)
 return (accuracy)
 })
plot <-plot(cost_values, accuracy_values, type = "b")
print(plot)

jpeg("svm_plot.jpg")
plot <-plot(cost_values, accuracy_values, type = "b")
dev.off()

```



```{r}
set.seed(12345)
svm_model <- ksvm(Letter ~ ., data = letters_train,kernel = "rbfdot",C=12)
svm_pred <- predict(svm_model, letters_test)

agreement_C12 <- svm_pred == letters_test$Letter
table(agreement_C12)

prop.table(table(agreement_C12))

```




